---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# routineqc

<!-- badges: start -->
<!-- badges: end -->

The goal of routineqc is to provide reusable quality-control tools for routine
health facility-month surveillance data. It is designed to help identify likely
data-entry anomalies while preserving raw records and appending QC flags.

## Status: experimental (0.1.0)

`routineqc` 0.1.0 is an **experimental release candidate**. APIs, QC
thresholds, and the action policy may still change before a stable release.
Treat every output as an aid to review, not a validated production tool.

- **Anomaly flags are evidence for review, not automatic proof of a
  data-entry error.** Only impossible core tested/positive counts are
  auto-excluded under the default `conservative_review` policy; attendance,
  statistical, temporal, and tested-volume signals remain review-only.
- The prevalence model, tested-volume baselines, and action policy have been
  validated against synthetic fixtures only; no upstream production pipeline
  has yet been integrated (see `MIGRATION.md`).
- The Run Explorer is read-only: there is no reviewer decision, annotation, or
  audit-trail data model yet.
- Saved QC runs (the `.rds` file and its manifest) contain the complete
  flagged dataset, including facility identifiers, and must be treated as
  confidential, equivalent to source data. The fingerprint recorded in
  provenance is an identifier, not anonymization.

See `docs/ACTION_POLICY.md`, `docs/TESTED_VOLUME_BASELINES.md`,
`docs/PREVALENCE_MODEL.md`, and `docs/RUN_EXPLORER.md` for full policy, model,
and privacy detail.

## Quick start

From the package directory, run this in an R terminal:

``` r
# install.packages('pak')
pak::pkg_install('.')
system2('Rscript', c('scripts/smoke_test_routineqc.R', '--launch'))
```

The smoke test builds a synthetic QC run, checks that the package can complete
the main workflow, and opens the local QC Run Explorer when `--launch` is
included. The app listens on a localhost URL; keep the R terminal open while you
review it in the browser, then stop the process when finished.

## Installation

Install from source in this repository:

``` r
# install.packages("pak")
pak::pkg_install("path/to/routineqc")
```

### Installing this exact release candidate

External testers exercising a specific tagged release (for example, a
reviewer following up on the 0.1.0 release candidate) should install from
that tag rather than an untagged checkout, so results are reproducible:

``` r
# git clone https://github.com/jt-hicks/routineqc.git
# cd routineqc && git checkout v0.1.0

# install.packages('pak')
pak::pkg_install('.')
```

Then confirm the install and exercise the package end to end with the
non-interactive smoke test, adding `--launch` to also open the local Run
Explorer:

``` r
system2('Rscript', c('scripts/smoke_test_routineqc.R', '--launch'))
```

Report any install or smoke-test failure against the exact commit or tag
checked out.

## Minimal simulation example

```{r simulation-example}
library(routineqc)

sim <- simulate_qc_data(n_facilities = 20, n_months = 24, seed = 42)
head(sim)
```

## Run routine QC

```{r run-qc}
qc <- suppressWarnings(run_routine_qc(
  raw_data = sim,
  facility_var = "facility",
  region_var = "region",
  council_var = "council",
  district_var = "district",
  month_var = "month",
  tested_var = "tested",
  positive_var = "positive",
  nthreads = 1
))

names(qc$summaries)
qc$summaries$by_region
qc$manifest$model_prediction_coverage
table(qc$data_flagged$prediction_status)
```

## Adapt pipeline data

Pipeline-specific names and conventions belong in an adapter, outside the QC
engine. A generic declarative adapter is included:

``` r
adapted <- adapt_qc_data(
  pipeline_data,
  mapping = list(
    facility_id = 'org_unit_id', month_date = 'reporting_month',
    region = 'province', tested = 'malaria_tested',
    positive = 'malaria_positive'
  ),
  identity_strategy = 'DHIS2 organisation unit ID',
  adapter = 'example_pipeline', adapter_version = '0.1.0',
  dataset_id = 'curated-2026-08', source_system = 'DHIS2'
)

adapted$diagnostics
qc <- run_qc_adapter(adapted)
```

Adapter errors block QC without deleting records. Count-conversion warnings do
not block QC: the converted `NA` values remain available to the logical checks.

The default `conservative_review` policy authorizes exclusion only for
impossible core tested/positive records. Attendance contradictions and
statistical, temporal, or tested-volume anomalies remain review-only.

Use a flags-only run when all exclusion decisions must remain external:

``` r
flags_only <- qc_config(action_policy = list(name = 'flags_only'))
```

Operational tested-volume baselines use only previous calendar months. For a
substantially complete historical dataset, explicitly select the retrospective
profile, which records that future observations were used:

``` r
historical <- qc_config('retrospective')
```

Each result is a validated, versioned run object with deterministic input and
analysis identities and a unique execution identity. Persist complete runs with
a metadata-only JSON manifest:

``` r
write_qc_run(qc, 'qc-runs/example.rds')
restored <- read_qc_run('qc-runs/example.rds')
```

The RDS contains the complete flagged dataset and must be protected like the
source data. The fingerprint is an identifier, not anonymization.

## Explore saved runs locally

The read-only Run Explorer discovers paired artifacts, validates a selected run,
and provides overview, review-queue, diagnostic, configuration, and provenance
views:

``` r
list_qc_runs('qc-runs')
launch_qc_app('qc-runs')
```

The app binds to localhost, performs no writes, and provides no record download.
Treat the local browser session and run directory as confidential.
The review queue offers focused column views with hover definitions. Facility
diagnostics compare observed prevalence with stored GAM expectations without
rerunning the model.

The Overview combines a mutually exclusive QC-action donut, an individual-reason
bar chart, and facility and district burden summaries. A row with multiple QC
reasons contributes once to each applicable reason bar, so reason counts need
not sum to the number of rows. The always-visible flow diagram explains how
evidence is mapped to retain, review, or authorized exclusion actions.

Interactive diagnostics expose row details on hover and compare observed
prevalence and testing volume with their stored expectations. The District tab
provides faceted facility time series, switchable between prevalence and number
tested, when the upstream adapter supplies district.

To exercise the complete package with synthetic data before connecting a real
pipeline, run from the package repository:

``` sh
Rscript scripts/smoke_test_routineqc.R
```

The smoke fixture contains three districts with eight facilities each and 12 to
24 months per facility. Pass an output directory as the first argument and add
`--launch` to open the generated run immediately.

routineqc preserves raw data and appends QC outputs as additional columns
rather than deleting records.

Anomaly flags are evidence for review, not automatic proof of error. See
`docs/ACTION_POLICY.md` and `docs/TESTED_VOLUME_BASELINES.md` for policy and
baseline limitations, `docs/PREVALENCE_MODEL.md` for model assessment semantics,
`docs/ADAPTERS.md` for the pipeline boundary, and `docs/RUN_ARTIFACTS.md` for
persistence guidance. See `docs/RUN_EXPLORER.md` for application scope and
privacy constraints.

## Notes

You can regenerate README.md from this file with:

``` r
devtools::build_readme()
```
