---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# routineqc

<!-- badges: start -->
<!-- badges: end -->

The goal of routineqc is to provide reusable quality-control tools for routine
health facility-month surveillance data. It is designed to help identify likely
data-entry anomalies while preserving raw records and appending QC flags.

## Installation

Install from source in this repository:

``` r
# install.packages("pak")
pak::pkg_install("path/to/routineqc")
```

## Minimal simulation example

```{r simulation-example}
library(routineqc)

sim <- simulate_qc_data(n_facilities = 20, n_months = 24, seed = 42)
head(sim)
```

## Run routine QC

```{r run-qc}
qc <- run_routine_qc(
  raw_data = sim,
  facility_var = "facility",
  region_var = "region",
  council_var = "council",
  district_var = "district",
  month_var = "month",
  tested_var = "tested",
  positive_var = "positive",
  nthreads = 1
)

names(qc$summaries)
qc$summaries$by_region
```

The default `conservative_review` policy authorizes exclusion only for
impossible core tested/positive records. Attendance contradictions and
statistical, temporal, or tested-volume anomalies remain review-only.

Use a flags-only run when all exclusion decisions must remain external:

``` r
flags_only <- qc_config(action_policy = list(name = 'flags_only'))
```

Operational tested-volume baselines use only previous calendar months. For a
substantially complete historical dataset, explicitly select the retrospective
profile, which records that future observations were used:

``` r
historical <- qc_config('retrospective')
```

routineqc preserves raw data and appends QC outputs as additional columns
rather than deleting records.

Anomaly flags are evidence for review, not automatic proof of error. See
`docs/ACTION_POLICY.md` and `docs/TESTED_VOLUME_BASELINES.md` for policy and
baseline limitations.

## Notes

You can regenerate README.md from this file with:

``` r
devtools::build_readme()
```
