---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# routineqc

<!-- badges: start -->
<!-- badges: end -->

The goal of routineqc is to provide reusable quality-control tools for routine
health facility-month surveillance data. It is designed to help identify likely
data-entry anomalies while preserving raw records and appending QC flags.

## Installation

Install from source in this repository:

``` r
# install.packages("pak")
pak::pkg_install("path/to/routineqc")
```

## Minimal simulation example

```{r simulation-example}
library(routineqc)

sim <- simulate_qc_data(n_facilities = 20, n_months = 24, seed = 42)
head(sim)
```

## Run routine QC

```{r run-qc}
qc <- run_routine_qc(
  raw_data = sim,
  facility_var = "facility",
  region_var = "region",
  council_var = "council",
  district_var = "district",
  month_var = "month",
  tested_var = "tested",
  positive_var = "positive",
  nthreads = 1
)

names(qc$summaries)
qc$summaries$by_region
qc$manifest$model_prediction_coverage
table(qc$data_flagged$prediction_status)
```

## Adapt pipeline data

Pipeline-specific names and conventions belong in an adapter, outside the QC
engine. A generic declarative adapter is included:

``` r
adapted <- adapt_qc_data(
  pipeline_data,
  mapping = list(
    facility_id = 'org_unit_id', month_date = 'reporting_month',
    region = 'province', tested = 'malaria_tested',
    positive = 'malaria_positive'
  ),
  identity_strategy = 'DHIS2 organisation unit ID',
  adapter = 'example_pipeline', adapter_version = '0.1.0',
  dataset_id = 'curated-2026-08', source_system = 'DHIS2'
)

adapted$diagnostics
qc <- run_qc_adapter(adapted)
```

Adapter errors block QC without deleting records. Count-conversion warnings do
not block QC: the converted `NA` values remain available to the logical checks.

The default `conservative_review` policy authorizes exclusion only for
impossible core tested/positive records. Attendance contradictions and
statistical, temporal, or tested-volume anomalies remain review-only.

Use a flags-only run when all exclusion decisions must remain external:

``` r
flags_only <- qc_config(action_policy = list(name = 'flags_only'))
```

Operational tested-volume baselines use only previous calendar months. For a
substantially complete historical dataset, explicitly select the retrospective
profile, which records that future observations were used:

``` r
historical <- qc_config('retrospective')
```

Each result is a validated, versioned run object with deterministic input and
analysis identities and a unique execution identity. Persist complete runs with
a metadata-only JSON manifest:

``` r
write_qc_run(qc, 'qc-runs/example.rds')
restored <- read_qc_run('qc-runs/example.rds')
```

The RDS contains the complete flagged dataset and must be protected like the
source data. The fingerprint is an identifier, not anonymization.

## Explore saved runs locally

The read-only Run Explorer discovers paired artifacts, validates a selected run,
and provides overview, review-queue, diagnostic, configuration, and provenance
views:

``` r
list_qc_runs('qc-runs')
launch_qc_app('qc-runs')
```

The app binds to localhost, performs no writes, and provides no record download.
Treat the local browser session and run directory as confidential.

routineqc preserves raw data and appends QC outputs as additional columns
rather than deleting records.

Anomaly flags are evidence for review, not automatic proof of error. See
`docs/ACTION_POLICY.md` and `docs/TESTED_VOLUME_BASELINES.md` for policy and
baseline limitations, `docs/PREVALENCE_MODEL.md` for model assessment semantics,
`docs/ADAPTERS.md` for the pipeline boundary, and `docs/RUN_ARTIFACTS.md` for
persistence guidance. See `docs/RUN_EXPLORER.md` for application scope and
privacy constraints.

## Notes

You can regenerate README.md from this file with:

``` r
devtools::build_readme()
```
