---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# routineqc

<!-- badges: start -->
<!-- badges: end -->

The goal of routineqc is to provide reusable quality-control tools for routine
health facility-month surveillance data. It is designed to help identify likely
data-entry anomalies while preserving raw records and appending QC flags.

## Installation

Install from source in this repository:

``` r
# install.packages("pak")
pak::pkg_install("path/to/routineqc")
```

## Minimal simulation example

```{r simulation-example}
library(routineqc)

sim <- simulate_qc_data(n_facilities = 20, n_months = 24, seed = 42)
head(sim)
```

## Run routine QC

```{r run-qc}
qc <- run_routine_qc(
  raw_data = sim,
  facility_var = "facility",
  region_var = "region",
  council_var = "council",
  district_var = "district",
  month_var = "month",
  tested_var = "tested",
  positive_var = "positive",
  nthreads = 1
)

names(qc$summaries)
qc$summaries$by_region
```

routineqc preserves raw data and appends QC outputs as additional columns
rather than deleting records.

Flagged records should be interpreted as records for review and sensitivity
analysis, not automatic proof of error.

## Notes

You can regenerate README.md from this file with:

``` r
devtools::build_readme()
```
