---
title: 'Adapting a Pipeline Extract for routineqc'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adapting a Pipeline Extract for routineqc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
```

This example uses synthetic data. An adapter translates source conventions but
does not delete, aggregate, impute, or make QC action decisions.

```{r}
library(routineqc)

source_data <- simulate_qc_data(n_facilities = 12, n_months = 18, seed = 42)
source_data$record_id <- sprintf('synthetic-%04d', seq_len(nrow(source_data)))

adapted <- adapt_qc_data(
  source_data,
  mapping = list(
    facility_id = 'facility',
    month_date = 'month',
    region = 'region',
    district = 'district',
    council = 'council',
    tested = 'tested',
    positive = 'positive',
    source_record_id = 'record_id'
  ),
  identity_strategy = 'synthetic stable facility code',
  adapter = 'synthetic_vignette',
  adapter_version = '1.0.0',
  dataset_id = 'synthetic-seed-42',
  source_system = 'routineqc-simulator'
)

adapted$ready
head(adapted$diagnostics)
```

Blocking diagnostics must be resolved upstream without dropping records. A
ready result can be passed directly into the QC engine.

```{r}
qc_run <- suppressWarnings(run_qc_adapter(adapted, nthreads = 1))

qc_run$manifest$provenance
qc_run$manifest$model_prediction_coverage
qc_run$summaries$by_region
```

The resulting object has the same versioned configuration, provenance,
integrity checks, model coverage, flags, and action policy as a direct
`run_routine_qc()` call.

## Optional attendance data

Attendance checks are applied only when the source supplies an approved,
documented attendance denominator. If `attending` is mapped, the adapter also
requires a non-empty `attendance_definition`:

```{r, eval = FALSE}
adapted_with_attendance <- adapt_qc_data(
  source_data,
  mapping = list(
    facility_id = 'facility_code', month_date = 'reporting_month',
    region = 'region', tested = 'tested', positive = 'positive',
    attending = 'eligible_anc_attendees'
  ),
  attendance_definition = paste(
    'ANC attendees eligible for malaria testing, as defined in the',
    'approved reporting specification'
  ),
  identity_strategy = 'stable facility code',
  adapter = 'example_pipeline', adapter_version = '1.0.0',
  dataset_id = 'example-extract', source_system = 'source-system'
)
```

Do not map a convenient attendance-like field unless its meaning and
comparability with `tested` have been approved. Data without attendance, or
with an entirely missing attendance column, can still be processed without
attendance checks.

## Action policy

The default `conservative_review` policy authorizes exclusion only for
impossible or missing tested/positive counts. Other anomalies remain evidence
for review. To keep every decision external to the package, pass a flags-only
configuration:

```{r, eval = FALSE}
flags_only <- qc_config(action_policy = list(name = 'flags_only'))
qc_run <- run_qc_adapter(adapted, config = flags_only)
```
