---
title: 'Exploring Persisted QC Runs'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring Persisted QC Runs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
```

The Run Explorer consumes persisted, validated QC artifacts. This synthetic
example creates one such artifact without using confidential records.

```{r}
library(routineqc)

source_data <- simulate_qc_data(n_facilities = 8, n_months = 12, seed = 7)
qc_run <- suppressWarnings(run_routine_qc(
  source_data,
  facility_var = 'facility', region_var = 'region', month_var = 'month',
  tested_var = 'tested', positive_var = 'positive',
  district_var = 'district', council_var = 'council',
  provenance = list(dataset_id = 'run-explorer-vignette'), nthreads = 1
))

run_dir <- tempfile('routineqc-runs-')
dir.create(run_dir)
write_qc_run(qc_run, file.path(run_dir, 'synthetic-run.rds'))
list_qc_runs(run_dir)
```

In an interactive local R session, launch the read-only application with:

```{r, eval = FALSE}
launch_qc_app(run_dir)
```

The browser can filter the review queue and display diagnostic plots, but it
cannot modify or download the stored records. The selected RDS and adjacent JSON
manifest must agree before the run is displayed.

```{r, include = FALSE}
unlink(run_dir, recursive = TRUE)
```
