---
title: 'Exploring Persisted QC Runs'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring Persisted QC Runs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
```

The Run Explorer consumes persisted, validated QC artifacts. This synthetic
example creates one such artifact without using confidential records.

```{r}
library(routineqc)

source_data <- simulate_qc_data(n_facilities = 8, n_months = 12, seed = 7)
qc_run <- suppressWarnings(run_routine_qc(
  source_data,
  facility_var = 'facility', region_var = 'region', month_var = 'month',
  tested_var = 'tested', positive_var = 'positive',
  district_var = 'district', council_var = 'council',
  provenance = list(dataset_id = 'run-explorer-vignette'), nthreads = 1
))

run_dir <- tempfile('routineqc-runs-')
dir.create(run_dir)
write_qc_run(qc_run, file.path(run_dir, 'synthetic-run.rds'))
list_qc_runs(run_dir)
```

In an interactive local R session, launch the read-only application with:

```{r, eval = FALSE}
launch_qc_app(run_dir)
```

The browser is organized around four review tasks:

- **Overview:** inspect the mutually exclusive QC-action donut, individual QC
  reason counts, facility and district burden, and the QC action flow. A row
  with several reasons contributes once to every applicable reason bar.
- **Review queue:** filter flagged records and switch among focused column sets.
  Hover over a column heading to see its definition.
- **Diagnostics:** compare facility observations with stored prevalence-model
  expectations and tested-volume baselines. Plotly hover text exposes the row
  context used to interpret each point.
- **District:** inspect faceted facility time series for either prevalence or
  number tested. This view is available only when district was supplied by the
  upstream adapter.

The application cannot modify or download stored records, and it does not refit
models. The selected RDS and adjacent JSON manifest must agree before the run is
displayed. It binds to localhost, but both the run directory and browser session
must still be handled according to the source data's confidentiality.

```{r, include = FALSE}
unlink(run_dir, recursive = TRUE)
```
