---
title: "Routine Surveillance QC Report"
format:
  html:
    toc: true
    code-fold: true
execute:
  echo: false
---

```{r}
library(routineqc)
library(dplyr)
library(ggplot2)

# Expected object names in environment:
# raw_data: raw facility-month dataset
# qc_result: output from run_routine_qc(raw_data, ...)
```

## Dataset Overview

```{r}
if (!exists("qc_result")) {
  stop("Please provide qc_result <- run_routine_qc(...) before rendering report.")
}

cat("Rows:", nrow(qc_result$data_flagged), "\n")
cat("Facilities:", dplyr::n_distinct(qc_result$data_flagged$facility_id), "\n")
cat("Months:", dplyr::n_distinct(qc_result$data_flagged$month_date), "\n")
cat("Regions:", dplyr::n_distinct(qc_result$data_flagged$region), "\n")
```

## Run Configuration

```{r}
cat('Configuration profile:', qc_result$config$profile, '\n')
cat('Configuration schema:', qc_result$config$schema_version, '\n')
cat('Tested-volume baseline:', qc_result$config$tested_volume$baseline_mode, '\n')
cat('Baseline uses future reports:', qc_result$config$tested_volume$baseline_uses_future, '\n')
```

The trailing operational baseline has a warm-up period, can adapt slowly after
a genuine structural change, and may be contaminated by erroneous earlier
reports. The centered retrospective baseline uses future information, can change
when later reports are appended, and must not be interpreted as a real-time
alert. Missing calendar months reduce baseline evidence in both modes.

## QC Protocol Summary

The routineqc workflow applies:

1. Data preparation and standardisation.
2. Logical consistency checks for tested/positive counts.
3. GAM-based prevalence expectation modeling.
4. Residual and prevalence anomaly flags.
5. Tested volume anomaly flags from rolling median/MAD.
6. Temporal isolation checks for prevalence extremes.
7. Hierarchical action assignment.

## Logical Errors

```{r}
qc_result$data_flagged %>%
  summarise(
    invalid_logical_rows = sum(flag_invalid_logical, na.rm = TRUE),
    pct_invalid = 100 * mean(flag_invalid_logical, na.rm = TRUE)
  )
```

## Prevalence Anomalies

```{r}
qc_result$data_flagged %>%
  summarise(
    resid_extreme = sum(flag_resid_extreme, na.rm = TRUE),
    prevalence_extreme = sum(flag_prevalence_extreme, na.rm = TRUE),
    isolated_extreme = sum(flag_isolated_extreme, na.rm = TRUE)
  )

plot_prevalence_residuals(qc_result$data_flagged)
```

## Tested Volume Anomalies

```{r}
qc_result$data_flagged %>%
  summarise(
    tested_volume_extreme = sum(flag_tested_volume_extreme, na.rm = TRUE),
    baseline_not_ready = sum(!tested_baseline_ready, na.rm = TRUE),
    tested_jump = sum(flag_tested_large_jump, na.rm = TRUE),
    tested_drop = sum(flag_tested_large_drop, na.rm = TRUE)
  )

plot_tested_robust_z(qc_result$data_flagged)
```

## Combined QC Actions

```{r}
qc_result$data_flagged %>%
  count(qc_action, sort = TRUE)
```

## Region QC Summary

```{r}
qc_result$summaries$by_region
plot_qc_summary_by_region(qc_result$summaries$by_region)
```

## Council QC Summary

```{r}
if ("by_council" %in% names(qc_result$summaries)) {
  qc_result$summaries$by_council
} else {
  cat("Council field not available in this dataset.")
}
```

## District QC Summary

```{r}
if ("by_district" %in% names(qc_result$summaries)) {
  qc_result$summaries$by_district
} else {
  cat("District field not available in this dataset.")
}
```

## Before/After Summaries

```{r}
summarise_before_after_qc(qc_result$data_flagged, by = "region")
plot_before_after_prevalence(qc_result$data_flagged)
```

## Threshold Sensitivity

```{r}
run_threshold_sensitivity(qc_result$data_flagged, by = "region")
```

## Facility Audit

```{r}
qc_result$summaries$by_facility %>%
  arrange(desc(pct_any_qc_issue)) %>%
  head(20)
```

## Example Flagged Facility Plots

```{r}
flagged_facility <- qc_result$data_flagged %>%
  group_by(facility_id) %>%
  summarise(any_qc = any(flag_any_qc_issue, na.rm = TRUE), .groups = "drop") %>%
  filter(any_qc) %>%
  slice_head(n = 1) %>%
  pull(facility_id)

if (length(flagged_facility) == 1) {
  plot_facility_timeseries(qc_result$data_flagged, flagged_facility)
}
```
